Addition and removal energies of circular quantum dots

### 2017-05-10 Manuscript/paper.tex
Fixups

\newcommand{\D}{\operatorname{d\!}}
\newcommand{\E}{\operatorname{e}}
\newcommand{\bra}[1]{\langle #1 |}
\newcommand{\braket}[2]{\langle #1 | #2\rangle}
\newcommand{\ket}[1]{| #1 \rangle}
\newcommand{\normord}[1]{\mathopen: #1 \mathclose:}
\newcommand{\overbar}[1]{\mkern 3.0mu\overline{\mkern-3.0mu #1 \mkern-0.8mu}\mkern 0.8mu}

@@

\begin{figure}
\includegraphics{fig-diagrams-imsrg}
\caption{Diagrammatic representation of $D^{[k]}(\circ, \bullet)$ of \eqref{eq:flow0}, \eqref{eq:flow1}, and \eqref{eq:flow2}, where open circles represent $A$ and filled circles represent $B$.  The two-body vertices are implicitly antisymmetric (i.e.\ \textit{Hugenholtz diagrams} \cite{HUGENHOLTZ1957481}, as described in \cite{shavitt2009many} \S 4.4.3).  Internal directed lines (bound/dummy indices) that point upward denote summations over unoccupied states, whereas lines that point downward denote summations over occupied states.  External undirected lines denote unconstrained free indices.}
\label{fig:diagrams-imsrg}
\end{figure}

@@

Using the coupled cluster wave function, the Schr{\"o}dinger equation,
\begin{equation}\label{eq:cc_schrodeq}
\hat{H}e^{\hat{T}}\ket{\Phi} = Ee^{\hat{T}}\ket{\Phi},
\end{equation}
can be rewritten by left-multiplying by $\bra{\Phi}e^{-T}$ as,
\begin{equation}\label{eq:ccsd0}
\bra{\Phi}\overbar{H}_{CC}\ket{\Phi} = E,
\end{equation}
where we define an coupled cluster effective Hamiltonian,
\begin{equation}\label{eq:cc_heff0}
\overbar{H}_{CC} \equiv e^{-\hat{T}}\hat{H}e^{\hat{T}},
\end{equation}
in which the wave operator, $e^{\hat{T}}$, acts as a similarity transform on the Hamiltonian in the same way that $\hat{U}(s)$ acts to transform the Hamiltonian in SRG methods.  An important difference, however, is that the coupled cluster wave operator is not unitary, and thus $\overbar{H}$ is not Hermitian.

The effective Hamiltonian in \eqref{eq:cc_heff0} can be rewritten with commutators according to the Baker-Campbell-Hausdorff expansion as,
\begin{equation*}\label{eq:cc_bch}
\overbar{H}_{CC} = \hat{H} + [ \hat{H},\hat{T} ] + \frac{1}{2} [[ \hat{H},\hat{T} ],\hat{T} ] + \frac{1}{3!} [[[ \hat{H},\hat{T} ],\hat{T} ],\hat{T} ] + \frac{1}{4!} [[[[ \hat{H},\hat{T} ],\hat{T} ],\hat{T} ],\hat{T} ],
\end{equation*}
which terminates at four-nested commutators due to the two-body nature of the interaction.  Like with IM-SRG, this commutator expression ensures that coupled cluster is size-extensive and contains only connected terms.  In addition, because $\hat{T}$ is an excitation operator, terms of the form $\hat{T}\hat{H}$ are disconnected and thus vanish \cite{shavitt2009many}.  Therefore the coupled cluster effective Hamiltonian can be further reduced to
\begin{equation}\label{eq:cc_heff1}
\overbar{H}_{CC} = \left( \hat{H}e^{\hat{T}} \right)_{c},
\end{equation}
where the label ``$c$'' indicates that only connected terms are used.

@@

\begin{figure}
\includegraphics{fig-diagrams-ccsd}
\caption{Diagrammatic representation of $\overbar{H}$ of \eqref{eq:cc_heff1}, where open circles represent the excitation cluster operators $\hat{T}_{1}$ and $\hat{T}_{2}$, and filled circles represent the two-body interaction $\hat{H}^{[2]}$.  The diagrams are implicitly antisymmetrized (Hugenholtz diagrams).  Lines connected to $\hat{T}$ are always directed upward because they represent an excitation operator while the directions of external lines connected to $\hat{H}^{[2]}$ are unconstrained.}
\label{fig:diagrams-ccsd}
\end{figure}

The unknown cluster amplitudes in CCSD, $t_{a, i}^{[1]}$ and $t_{a, b, i, j}^{[2]}$, are calculated by left-multiplying \eqref{eq:cc_schrodeq} by $\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{a}}e^{-\hat{T}}$ and $\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{j}^\dagger \hat a_{b} \hat a_{a}}e^{-\hat{T}}$, respectively.
\begin{gather}\label{eq:ccsd1}
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{a}}\overbar{H}_{CC}\ket{\Phi} = 0 \\
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{j}^\dagger \hat a_{b} \hat a_{a}}\overbar{H}_{CC}\ket{\Phi} = 0 \notag
\end{gather}
After the Fock matrix has been diagonalized, the diagonal components of \eqref{eq:ccsd1} can be separated and, after expanding the exponent in \eqref{eq:cc_heff1}, the non-vanishing terms of the CCSD amplitude equations become,
\begin{gather}\label{eq:ccsd2}
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{a}} \hat{H}^{[2]}( \hat{T}_{1} + \hat{T}_{2} + \hat{T}_{1}\hat{T}_{2} + \frac{1}{2}\hat{T}_{1}^{2} + \frac{1}{3!}\hat{T}_{1}^{3} )_{c} \ket{\Phi} = (H_{i,i}^{[1]} - H_{a,a}^{[1]})t_{a, i}^{[1]} \\
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{j}^\dagger \hat a_{b} \hat a_{a}} \hat{H}^{[2]}( 1 + \hat{T}_{1} + \hat{T}_{2} + \frac{1}{2}\hat{T}_{1}^{2} + \hat{T}_{1}\hat{T}_{2} + \frac{1}{2}\hat{T}_{2}^{2} + \frac{1}{3!}\hat{T}_{1}^{3} + \frac{1}{2}\hat{T}_{1}^{2}\hat{T}_{2} + \frac{1}{4!}\hat{T}_{1}^{4} )_{c} \ket{\Phi} \notag \\
\hspace{10cm} = (H_{i,i}^{[1]} + H_{j,j}^{[1]} - H_{a,a}^{[1]} - H_{b,b}^{[1]})t_{a, b, i, j}^{[2]} \notag
\end{gather}
These non-linear equations are solved using an iterative procedure that begins with an initial guess for the cluster amplitudes which are used to calculate the left-hand side of \eqref{eq:ccsd2}. Dividing by the denominator on the right-hand side gives updated amplitudes which are used to reevaluate the left-hand side. This proccess is repeated until the amplitudes reach a fixed point. Like the Hartree-Fock iterative procedure, employing similar convergence acceleration techniques can help improve the coupled cluster iterations.

\subsubsection{EOM-CC}
\label{subsec:eomcc}
Like EOM-IM-SRG, the equation-of-motion technique can be applied after a coupled-cluster ground-state calculation, by using the coupled-cluster effective Hamiltonian.  Here, the non-Hermitian nature of $\overbar{H}_{CC}$ becomes apparent.  In this case, in addition to constructing excitation ladder operators \eqref{eq:gen_attached} and \eqref{eq:gen_removed} that correspond to the right-eigenvectors of the generalized eigenvalue problem \eqref{eq:EOMIMSRG}, there exist analogous de-excitation ladder operators, $\hat{L}^{(N+1)}$ and $\hat{L}^{(N-1)}$, that correspond to the left-eigenvectors,
\begin{align}
  \label{eq:left_attached}
    \hat{L}^{(N+1)}_w &= \sum_p \bar{n}_p l^{[1,0]}_p  \hat{a}_p + \frac{1}{2} \sum_{p, q, r} \bar{n}_p \bar{n}_q n_r l^{[2,1]}_{p, q, r} \hat{a}^\dagger_r \hat{a}_q \hat{a}_p + \cdots,  \\
  \label{eq:left_removed}
    \hat{L}^{(N-1)}_w &= \sum_p n_p l_p^{[0,1]} \hat{a}^\dagger_p + \frac{1}{2} \sum_{p, q, r} n_p n_q \bar{n}_r  l^{[1,2]}_{p, q, r} \hat{a}^\dagger_p  \hat{a}^\dagger_q \hat{a}_r + \cdots.
\end{align}
These left-eigenvectors satisfy the left-eigenvalue problem, with left-eigenvalues $\varepsilon^{(\pm)}_w$, analogous to \eqref{eq:EOMIMSRG},
\begin{equation}\label{eq:left_EOM}
  \bra{\Phi^{(N)}_0} [\overbar{H},\bar{L}^{(N \pm 1)}_w] = \pm \varepsilon^{(\pm)}_w \bra{\Phi^{(N)}_0} \overbar{L}^{(N \pm 1)}_w,
\end{equation}
and form a bi-orthogonal set with the right-eigenvectors, $\braket{\bar{L}^{(N \pm 1)}_w}{\bar{X}^{(N \pm 1)}_u} = \delta_{w,u}$.

>>>>

\newcommand{\D}{\operatorname{d\!}}
\newcommand{\E}{\operatorname{e}}
\newcommand{\bra}[1]{\langle #1 |}
\newcommand{\braket}[2]{\langle #1 | #2\rangle}
\newcommand{\ket}[1]{| #1 \rangle}
\newcommand{\normord}[1]{\mathopen: #1 \mathclose:}

@@

\begin{figure}
\includegraphics{fig-diagrams-imsrg}
\caption{Diagrammatic representation of $D^{[k]}(\circ, \bullet)$ of \eqref{eq:flow0}, \eqref{eq:flow1}, and \eqref{eq:flow2}, where open circles represent $A$ and filled circles represent $B$.  The two-body vertices are implicitly antisymmetric (i.e.\ \textit{Hugenholtz diagrams} \cite{HUGENHOLTZ1957481}, as described in \cite{shavitt2009many} \S 4.4.3).  Internal directed lines (bound/dummy indices) that point upward denote summations over unoccupied states, whereas lines that point downward denote summations over occupied states.  External undirected lines denote unconstrained free indices.  In IM-SRG(2), the three-body diagrams are not included.}
\label{fig:diagrams-imsrg}
\end{figure}

@@

Using the coupled cluster wave function, the Schr{\"o}dinger equation,
\begin{equation}\label{eq:cc_schrodeq}
\hat{H}e^{\hat{T}}\ket{\Phi} = Ee^{\hat{T}}\ket{\Phi},
\end{equation}
can be rewritten by left-multiplying by $\bra{\Phi}e^{-T}$ as,
\begin{equation}\label{eq:ccsd0}
\bra{\Phi}\bar{H}_{\mathrm{CC}}\ket{\Phi} = E,
\end{equation}
where we define an coupled cluster effective Hamiltonian,
\begin{equation}\label{eq:cc_heff0}
\bar{H}_{\mathrm{CC}} \equiv e^{-\hat{T}}\hat{H}e^{\hat{T}},
\end{equation}
in which the wave operator, $e^{\hat{T}}$, acts as a similarity transform on the Hamiltonian in the same way that $\hat{U}(s)$ acts to transform the Hamiltonian in SRG methods.  An important difference, however, is that the coupled cluster wave operator is not unitary, and thus $\bar{H}$ is not Hermitian.

The effective Hamiltonian in \eqref{eq:cc_heff0} can be rewritten with commutators according to the Baker-Campbell-Hausdorff expansion as,
\begin{equation*}\label{eq:cc_bch}
\bar{H}_{\mathrm{CC}} = \hat{H} + [ \hat{H},\hat{T} ] + \frac{1}{2} [[ \hat{H},\hat{T} ],\hat{T} ] + \frac{1}{3!} [[[ \hat{H},\hat{T} ],\hat{T} ],\hat{T} ] + \frac{1}{4!} [[[[ \hat{H},\hat{T} ],\hat{T} ],\hat{T} ],\hat{T} ],
\end{equation*}
which terminates at four-nested commutators due to the two-body nature of the interaction.  Like with IM-SRG, this commutator expression ensures that coupled cluster is size-extensive and contains only connected terms.  In addition, because $\hat{T}$ is an excitation operator, terms of the form $\hat{T}\hat{H}$ are disconnected and thus vanish \cite{shavitt2009many}.  Therefore the coupled cluster effective Hamiltonian can be further reduced to
\begin{equation}\label{eq:cc_heff1}
\bar{H}_{\mathrm{CC}} = \left( \hat{H}e^{\hat{T}} \right)_{\mathrm{c}},
\end{equation}
where the subscript ``$\mathrm{c}$'' indicates that only connected terms are used.

@@

\begin{figure}
\includegraphics{fig-diagrams-ccsd}
\caption{Diagrammatic representation of $\bar{H}$ of \eqref{eq:cc_heff1}, where open circles represent the excitation cluster operators $\hat{T}_{1}$ and $\hat{T}_{2}$, and filled circles represent the two-body interaction $\hat{H}^{[2]}$.  The diagrams are implicitly antisymmetrized (Hugenholtz diagrams).  Lines connected to $\hat{T}$ are always directed upward because they represent an excitation operator while the directions of external lines connected to $\hat{H}^{[2]}$ are unconstrained.}
\label{fig:diagrams-ccsd}
\end{figure}

The unknown cluster amplitudes in CCSD, $t_{a, i}^{[1]}$ and $t_{a, b, i, j}^{[2]}$, are calculated by left-multiplying \eqref{eq:cc_schrodeq} by $\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{a}}e^{-\hat{T}}$ and $\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{j}^\dagger \hat a_{b} \hat a_{a}}e^{-\hat{T}}$, respectively.
\begin{gather}\label{eq:ccsd1}
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{a}}\bar{H}_{\mathrm{CC}}\ket{\Phi} = 0 \\
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{j}^\dagger \hat a_{b} \hat a_{a}}\bar{H}_{\mathrm{CC}}\ket{\Phi} = 0 \notag
\end{gather}
After the Fock matrix has been diagonalized, the diagonal components of \eqref{eq:ccsd1} can be separated and, after expanding the exponent in \eqref{eq:cc_heff1}, the non-vanishing terms of the CCSD amplitude equations become,
\begin{gather}\label{eq:ccsd2}
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{a}} \hat{H}^{[2]}( \hat{T}_{1} + \hat{T}_{2} + \hat{T}_{1}\hat{T}_{2} + \frac{1}{2}\hat{T}_{1}^{2} + \frac{1}{3!}\hat{T}_{1}^{3} )_{c} \ket{\Phi} = (H_{i,i}^{[1]} - H_{a,a}^{[1]})t_{a, i}^{[1]} \\
\bra{\Phi}\normord{ \hat a_{i}^\dagger \hat a_{j}^\dagger \hat a_{b} \hat a_{a}} \hat{H}^{[2]}( 1 + \hat{T}_{1} + \hat{T}_{2} + \frac{1}{2}\hat{T}_{1}^{2} + \hat{T}_{1}\hat{T}_{2} + \frac{1}{2}\hat{T}_{2}^{2} + \frac{1}{3!}\hat{T}_{1}^{3} + \frac{1}{2}\hat{T}_{1}^{2}\hat{T}_{2} + \frac{1}{4!}\hat{T}_{1}^{4} )_{c} \ket{\Phi} \notag \\
\hspace{10cm} = (H_{i,i}^{[1]} + H_{j,j}^{[1]} - H_{a,a}^{[1]} - H_{b,b}^{[1]})t_{a, b, i, j}^{[2]} \notag
\end{gather}
These non-linear equations are solved using an iterative procedure that begins with an initial guess for the cluster amplitudes which are used to calculate the left-hand side of \eqref{eq:ccsd2}. Dividing by the denominator on the right-hand side gives updated amplitudes which are used to reevaluate the left-hand side. This process is repeated until the amplitudes reach a fixed point. Like the Hartree-Fock iterative procedure, employing similar convergence acceleration techniques can help improve the coupled cluster iterations.

\subsubsection{EOM-CC}
\label{subsec:eomcc}
Like EOM-IM-SRG, the equation-of-motion technique can be applied after a coupled-cluster ground-state calculation, by using the coupled-cluster effective Hamiltonian.  Here, the non-Hermitian nature of $\bar{H}_{\mathrm{CC}}$ becomes apparent.  In this case, in addition to constructing excitation ladder operators \eqref{eq:gen_attached} and \eqref{eq:gen_removed} that correspond to the right-eigenvectors of the generalized eigenvalue problem \eqref{eq:EOMIMSRG}, there exist analogous de-excitation ladder operators, $\hat{L}^{(N+1)}$ and $\hat{L}^{(N-1)}$, that correspond to the left-eigenvectors,
\begin{align}
  \label{eq:left_attached}
    \hat{L}^{(N+1)}_w &= \sum_p \bar{n}_p l^{[1,0]}_p  \hat{a}_p + \frac{1}{2} \sum_{p, q, r} \bar{n}_p \bar{n}_q n_r l^{[2,1]}_{p, q, r} \hat{a}^\dagger_r \hat{a}_q \hat{a}_p + \cdots,  \\
  \label{eq:left_removed}
    \hat{L}^{(N-1)}_w &= \sum_p n_p l_p^{[0,1]} \hat{a}^\dagger_p + \frac{1}{2} \sum_{p, q, r} n_p n_q \bar{n}_r  l^{[1,2]}_{p, q, r} \hat{a}^\dagger_p  \hat{a}^\dagger_q \hat{a}_r + \cdots.
\end{align}
These left-eigenvectors satisfy the left-eigenvalue problem, with left-eigenvalues $\varepsilon^{(\pm)}_w$, analogous to \eqref{eq:EOMIMSRG},
\begin{equation}\label{eq:left_EOM}
  \bra{\Phi^{(N)}_0} [\bar{H},\bar{L}^{(N \pm 1)}_w] = \pm \varepsilon^{(\pm)}_w \bra{\Phi^{(N)}_0} \bar{L}^{(N \pm 1)}_w,
\end{equation}
and form a bi-orthogonal set with the right-eigenvectors, $\braket{\bar{L}^{(N \pm 1)}_w}{\bar{X}^{(N \pm 1)}_u} = \delta_{w,u}$.
